---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# string2path

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

This is an experimental R package using some Rust code to extract path information from TTF font file.

## Used Rust crates

* [rusttype](https://gitlab.redox-os.org/redox-os/rusttype/): For reading TTF data.
* [lyon](https://github.com/nical/lyon/): For tessellation of polygons and flattening the curves.

## Installation

If you are using Windows, you are lucky. Because this repository provides pre-compiled
binary for you, you don't need to install Rust toolchain.

Otherwise, you need to have Rust toolchain installed before trying to install this package. 
See <https://www.rust-lang.org/tools/install> for the installation instructions. 

``` r
# install.packages("devtools")
devtools::install_github("yutannihilation/string2path")
```

## Example

### `string2path()`

```{r example}
library(string2path)
library(ggplot2)

# This TTF file is downloaded from https://ipafont.ipa.go.jp/.
# For installed fonts, you can use systemfonts::system_fonts()
# to lookup the path.
d <- string2path("地獄お", "./fonts/ipam.ttf")

d <- tibble::rowid_to_column(d)

ggplot(d) +
  geom_path(aes(x, y, group = id, colour = factor(glyph_id))) +
  theme_minimal() +
  coord_equal() +
  theme(legend.position = "top")

library(gganimate)
d <- string2path("蹴", "./fonts/ipam.ttf")
d <- tibble::rowid_to_column(d)

ggplot(d) +
  geom_path(aes(x, y, group = id), size = 2, colour = "purple2", lineend = "round") +
  theme_minimal() +
  coord_equal() +
  transition_reveal(rowid)
```
### `string2fill()`

```{r example2}
# Sorry for my laziness, please replace the font path to the appropriate location in your system...
ttf_file <- "/usr/share/fonts/TTF/iosevka-heavyitalic.ttf"

d <- string2fill("abc", ttf_file)

ggplot(d) +
  geom_polygon(aes(x, y, group = id, fill = factor(id %% 3)), colour = "grey", size = 0.1) +
  theme_minimal() +
  coord_equal() +
  theme(legend.position = "none") +
  # colors are derived from https://colorhunt.co/palette/207313
  scale_fill_manual(values = c("#ff4b5c", "#056674", "#66bfbf"))
```

### `string2stroke()`

```{r string2stroke,animation.hook='gifski'}
for (w in 1:9 / 4) {
  d <- string2stroke("abc", ttf_file, line_width = w)
  
  p <- ggplot(d) +
    geom_polygon(aes(x, y, group = id, fill = factor(id %% 2)), colour = "grey", size = 0.1) +
    theme_minimal() +
    coord_equal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = c("purple", "pink"))
  plot(p)
}
```



## `tolerance`

`tolerance` controls resolution of the tessellation. You can reduce tolerance to get higher resolutions.

```{r example3, animation.hook='gifski'}
for (tolerance in c(0.5, 0.1, 0.05, 0.01, 0.005, 0.001, 0.0001, 0.00001)) {
  d <- string2fill("abc", ttf_file, tolerance = tolerance)
  
  p <- ggplot(d) +
    geom_polygon(aes(x, y, group = id), fill = "transparent", colour = "black", size = 0.5) +
    theme_minimal() +
    coord_equal() +
    ggtitle(paste0("tolerance: ", tolerance))
  plot(p)
}
```

Note that `tolerance` parameter behaves differently depending on the types of the result. Maybe this is somehow related to
whether intersection is allowed or not? I'm not sure...

```{r example4, animation.hook='gifski'}
for (tolerance in c(0.5, 0.1, 0.05, 0.01, 0.005, 0.001, 0.0001, 0.00001)) {
  d <- string2path("abc", ttf_file, tolerance = tolerance)
  
  p <- ggplot(d) +
    geom_path(aes(x, y, group = id), colour = "black", size = 0.5) +
    geom_point(aes(x, y, group = id), colour = "black", size = 1.5) +
    theme_minimal() +
    coord_equal() +
    ggtitle(paste0("tolerance: ", tolerance))
  plot(p)
}
```

## Resources

If you are curious about how you can use R and Rust, the following resources might help:

* [Using Rust code in R packages](https://jeroen.github.io/erum2018) by Jeroen Ooms
* <https://github.com/r-rust/hellorust/> is useful to see the basic structure
* <https://github.com/clauswilke/sinab> is also a cool R package with the power of Rust.
* ["Calling Rust code from C" section of Rustonomicon](https://doc.rust-lang.org/nomicon/ffi.html?highlight=extern#calling-rust-code-from-c). This page describes mostly about how to call C code from Rust, but it's useful to know how it works as the vice versa is very similar.
* [An answer to the question "How to return byte array from Rust function to FFI C?"](https://users.rust-lang.org/t/how-to-return-byte-array-from-rust-function-to-ffi-c/18136/4). This might be very obvious to those who are familiar with C, but it takes some time for me to figure out that I need to pass the data and the length at the same time, otherwise C side cannot know the size of the data.
